---
title: "Introduction and Motivation"
output: html_notebook
---
# Q1. What proportion of the p-values is below 0.05? *0.041*
```{r}
set.seed(1)
library(downloader)
url = "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/femaleControlsPopulation.csv"
filename = "femaleControlsPopulation.csv"
if (!file.exists(filename)) download(url,destfile=filename)
population = read.csv(filename)
pvals <- replicate(1000,{
  control = sample(population[,1],12)
  treatment = sample(population[,1],12)
  t.test(treatment,control)$p.val
})
head(pvals)
hist(pvals)
```
```{r}
mean(pvals < 0.05)
```

# Q2. What proportion of the p-values are below 0.01? *0.008*
```{r}
mean(pvals < 0.01)
```

# Q3. Assume you are testing the effectiveness of 20 diets on mice weight. For each of the 20 diets, you run an experiment with 10 control mice and 10 treated mice. Assume the null hypothesis that the diet has no effect is true for all 20 diets and that mice weights follow a normal distribution with mean 30 grams and a standard deviation of 2 grams. Run a Monte Carlo simulation for one of these studies:
```{r}
cases = rnorm(10,30,2)
controls = rnorm(10,30,2)
t.test(cases,controls)$p.value
```
Now run a Monte Carlo simulation imitating the results for the experiment for all 20 diets. If you set the seed at 100, set.seed(100), and use the same code as above inside a call to replicate(), how many of the p-values (number not proportion) are below 0.05? *1*
```{r}
set.seed(100)
pvals <- replicate(20,{
  cases = rnorm(10,30,2)
  controls = rnorm(10,30,2)
  t.test(cases,controls)$p.val
})
sum(pvals < 0.05)
```

# Q4. Now create a simulation to learn about the distribution of the number of p-values that are less than 0.05. In the previous question, we ran the 20 diet experiment once. Now we will run these 20 experiments 1,000 times and each time save the number of p-values that are less than 0.05.

Set the seed at 100 again, set.seed(100), run the code from the previous question 1,000 times, and save the number of times the p-value is less than 0.05 for each of the 1,000 instances.

What is the average of these 1,000 numbers? Note that this is the expected number of tests (out of the 20 we run) that we will reject when the null is true. *0.961*
```{r}
set.seed(100)
result <- function(trial) {
  pvals.list <-vector()
  replicate(trial, {
    pvals <- replicate(20,{
      cases = rnorm(10,30,2)
      controls = rnorm(10,30,2)
      t.test(cases,controls)$p.val
      })
    pvals.list <- append(pvals.list, sum(pvals < 0.05))
  })
}
mean(result(1000))
```

#Q5. Note that what the answer to question #4 says is that on average, we expect some p-value to be 0.05 even when the null is true for all diets.

Using the same simulation data from the question above, for what proportion of the 1,000 replicates do we reject the null hypothesis at least once (more than 0 false positives)?
```{r}
mean(result(1000) > 0)
```

